### 2024-02-16 paper/paper.qmd
Update scripts and paper qmd

---
title: "Tax-Equity-in-Low-and-Middle-Income-Countries-Replication"
author: 
  - Amie Liu, Wanling Ma
thanks: "Code and data are available at: https://github.com/WanlingMa/Tax-Equity-in-Low-and-Middle-Income-Countries-Replication"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Source

The original paper, "Tax Equity in Low- and Middle-Income Countries," utilized for replication, was published in the American Economic Association journal in 2024 [@article]. It explores the role of taxation in mitigating income inequality within countries with lower and middle incomes [@article]. This paper seeks to build upon the findings of the original paper, using the same datasets that were last retrieved on February 10th, 2024.

The two main research questions that were explored in this paper: TODO

## Methodology

This paper used the R programming language [@citeR] and the 'tidyverse' [@citeTidyverse] packages for replication, reproduction, data cleaning, and analysis. Within the 'tidyverse' package [@citeTidyverse], we used 'readr' [@citeReadr], 'dplyr' [@citeDplyr], 'tidyr' [@citeTidyr], and 'ggplot2' [@citeGgplot2] for reading CSV files, manipulating data, tidying data, and creating graphs.

Data collection methods exhibit significant diversity among open sources. We conducted a data-cleaning procedure for reproduction, focusing on variable selection, simplifying variable names, and removing missing values. The variables selected from the original paper were chosen based on their relevance to our research question, relating to (TODO).

This paper analyzes the datasets listed below, which we have organized into three categories: Economic Development and Growth, Labor Market and Employment, and Taxation and Income Distribution.

### Economic Development and Growth

`data/gdp_populaPon_WDI.dta`

The dataset was obtained from the World Bank’s World Development Indicators @worldDevelopmentIndicators. It encompasses a comprehensive sample from 266 countries and spans from the year 1960 to 2020 [@worldDevelopmentIndicators]. Three key variables available within the database are GDP in 2010 USD (gdp_2010usd), GDP in current USD (gdp_currentusd), and Population (pop). GDP per capita (USD) is calculated as the ratio of gdp_2010usd to population, providing a measure of economic output per person adjusted for inflation [@worldDevelopmentIndicators]. Meanwhile, GDP per capita (current) is calculated by dividing gdp_currentusd by population, offering a snapshot of economic output per person in current currency values [@worldDevelopmentIndicators]. This enables researchers to analyze economic growth, trends, and developments over an extensive time across diverse geographical regions.

`data/ross_mahdavi.dta`

The dataset was obtained from the Harvard Dataverse @oilGasData. It encompasses information about oil and natural gas production of all countries from 1932 to 2014 [@oilGasData]. It contains six column variables, including the country ID (country), year, approximate value of oil and gas product in US dollars (natres_richesse), GDP worldwide (gdp_wid), GDP in constant 2014 US dollars (gdp_2014usd), and oil and gas product as percentage of GDP (oil_pct) [@oilGasData]. Covering the period from 1932 to 2014, this dataset provides a longitudinal perspective on oil and natural gas production trends with a global scope, providing valuable insights into the economic aspects of oil and natural gas production.

`data/globalETR_bfjz.dta`

The dataset was sourced from the paper titled “Globalization and Factor Income Taxation” [@incomeTaxation] and is accessible through the website associated with the paper [@incomeTaxationWebsite]. It provides data on economic indicators and taxation in each country from 1965 to 2018, including the net domestic product in constant 2019 US dollars, effective tax rate, tax revenue, total income tax, social contributions, property and wealth tax, and indirect tax [@incomeTaxationWebsite]. This allows the authors to comparatively analyze the effective tax rate and taxation policies that influence income distribution globally.

`data/country_frame.dta`

The dataset was obtained from the World Bank’s World Development Indicators @incomeClassification. It includes several columns containing key indicators related to economic, social, and environmental development. These columns cover variables such as country codes represented by both two-letter (iso2) and three-letter (iso3), country names (country_name), region names (regionname), and income level classifications (incomelevelname) [@incomeClassification]. The income level classification consists of five distinct variables: high income, low income, upper middle income, lower middle income, and unclassified [@incomeClassification]. These variables provide insights into the economic, social, and environmental characteristics of countries worldwide.

### Labor Market and Employment

`data/API_SL.EMP.SELF.ZS_DS2_en_csv_v2_5560396.csv`

The dataset was sourced from the International Labour Organization @selfEmployedWorkforce and contains information from 1960 to 2022. It includes essential variables such as country code, country name, indicator name, indicator code, and the percentage of total employment each year [@selfEmployedWorkforce]. By focusing on self-employment as a specific indicator type, the dataset enables the analysis of the proportion of self-employed people among the total workforce across different countries and over time.

### Taxation and Income Distribution

`data/regressions_output_central.dta`
`data/Country_informaPon.xlsx`

These regression datasets were obtained from the authors’ article “Informality, Consumption Taxes, and Redistribution” @informalConsumption. The column variables include a unique identifier for each country (country_code), the year in which the data was collected (year), the number of iterations performed during the regression analysis (iteration), regression coefficient (b), standard error of the regression coefficient (se), and the coefficient of determination (r2_adj) [@informalConsumption]. These were conducted for the regression analysis of informal shares in household consumption data.

`data/PIT_parameters_AJ.dta`

The dataset originates from Jensen’s journal article “Employment Structure and the Rise of the Modern Tax System” @PITThresholds. It consists of data spanning from the year 1870 to 2014 and covers 100 countries. It includes several column variables, such as country code value (country_code), country name, year, mean of GDP per capita (lg_gdppc), mean of the marginal income tax rate (mtr), and mean of the size of personal income tax (size_pit) [@PITThresholds]. These variables delve into the relationship between country-level exemption threshold, tax system evolution, and economic development across different countries and times.

`data/PIT_Top_Rates_2022.csv`

The dataset was collected and coded by the authors in 2023, providing information on the country-level top marginal tax rate of personal income tax for the year 2022 across various countries [@article]. This dataset provides insights into the taxation policies and structures related to personal income tax rates implemented by different countries in the specified year.

`data/PSPR_incidence_dirtax_2023.dta`

The dataset utilized in the World Bank Poverty and Shared Prosperity report Chapter 5 "Correcting Course" @directTaxation is sourced from the Commitment for Equity Institute and the World Bank. It was shared by the World Bank in response to the authors' request [@directTaxation]. The dataset encompasses various variables, including country and year of the CEQ project (ctry_year), country of the CEQ project (ctry_ceq), year of the CEQ project (year), the incidence of direct taxes across income deciles (in_pdi_dirtax), World Economic Outlook group classification (class_weo), and others [@directTaxation]. These variables show poverty, inequality, and shared prosperity metrics, enabling detailed analyses of economic and equity-related indicators across different countries and periods.

# Results

## Figure 1
```{r}
source("../scripts/99-replications-figure1.R")
```

## Figure 2
```{r}
source("../scripts/100-replications-figure2.R")
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

# References